---
title: "Reproducible Research Project I"
author: "Andrew Mauro"
date: "May 31, 2017"
output: html_document
---

1. Read and Process the Data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list =ls())
library(dplyr)
library(ggplot2)
library(data.table) 

activity <- fread("activity.csv")

#format date column
activity$date <- as.POSIXct(activity$date, format = "%Y-%m-%d")

#data set is read and processed for our analysis.

```

2. Histogram of the total steps per day

```{r}
#I was unclear on what the question was asking for so I created two plots - one with the total steps for each day and a histogram of steps per day.

#create table of dates where count of date corresponds to step count
daily <- activity %>% group_by(date) %>% summarise(daily = sum(steps, na.rm = TRUE))

dates <- unique(daily$date)

df <- data.frame()

for(i in dates){
        vec <- rep(i, daily[daily$date == i, "daily"])
        vec <- data.frame(vec)
        df <- rbind(df, vec)
}

df$date <- as.POSIXct(df$vec, origin = "1970-01-01")

ggplot(df, aes(x = factor(x = date))) +
        geom_histogram(stat = "count") +
        ggtitle("Count of Steps per Day") +
        xlab("Date") +
        ylab ("Steps")



#frequency histogram of daily step counts counts

ggplot(daily, aes(x = daily)) +
        geom_histogram(bins = 10) +
        ggtitle("Count of Steps per Day") +
        xlab("Number of Steps") +
        ylab ("Frequency")

     
```


3. Mean and median number of steps taken each day

```{r}
#Produce table that summarizes mean and median steps per day
daystat <- activity %>% group_by(date) %>% summarise(meanStep = mean(steps, na.rm = TRUE), medStep = median(steps, na.rm = TRUE))

daystat

#the median and mode number of steps per day is 0.

```

4. Time series plot of the average number of steps taken


```{r}

ggplot(daystat, aes(x = date, y = meanStep)) +
        geom_line() +
        ggtitle("Average Steps Over Measurement Period") +
        xlab("Observation Day") +
        ylab ("Average Steps")


```


5. The 5-minute interval that, on average, contains the maximum number of steps

```{r}
#Produce table that summarizes mean steps per interval arranged by mean steps
interval <- activity %>% group_by(interval) %>% summarise(meanStep = mean(steps, na.rm = TRUE)) %>% arrange(desc(meanStep))

interval
#interval 83 - 206 steps on average.

```

6. Code to describe and show a strategy for imputing missing data

Missing values will be imputed based on the average number of steps per interval, grouped by day of week

```{r}
#identify weekends and week days
activity$weekday <- weekdays(activity$date)

#create new column depicting average steps for each grouped weekday/interval combination
activity <- activity %>% group_by(weekday, interval) %>% mutate(meanStep = round(mean(steps, na.rm = TRUE),0))

#replace NAs with average steps for the respective grouped weekday/interval combination
activity[is.na(activity$steps), "steps"] <- activity[is.na(activity$steps), "meanStep"]

activity
```

7. Histogram of the total number of steps taken each day after missing values are imputed

```{r}
#I was unclear on what the question was asking for so I created two plots - one with the total steps for each day and a histogram of steps per day.

#create table of dates where count of date corresponds to step count
daily <- activity %>% group_by(date) %>% summarise(daily = sum(steps, na.rm = TRUE))

dates <- unique(daily$date)

df <- data.frame()

for(i in dates){
        vec <- rep(i, daily[daily$date == i, "daily"])
        vec <- data.frame(vec)
        df <- rbind(df, vec)
}

df$date <- as.POSIXct(df$vec, origin = "1970-01-01")

ggplot(df, aes(x = factor(x = date))) +
        geom_histogram(stat = "count") +
        ggtitle("Count of Steps per Day") +
        xlab("Date") +
        ylab ("Steps")



#frequency histogram of daily step counts counts

ggplot(daily, aes(x = daily)) +
        geom_histogram(bins = 10) +
        ggtitle("Count of Steps per Day") +
        xlab("Number of Steps") +
        ylab ("Frequency")

```

8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}

activity <- activity %>% 
        mutate(wkndStatus = ifelse(weekday %in% c("Saturday", "sunday"),"Weekend", "Weekday" )) %>%
        group_by(interval, wkndStatus) %>% 
        mutate(intervalMean_gbWknd = mean(steps))

ggplot(activity, aes(x = interval, y = intervalMean_gbWknd)) +
        geom_bar(stat = "identity") +
        facet_wrap(~wkndStatus) +
        ggtitle("Mean Steps per Interval: Weekends vs. Weekdays") +
        xlab("Interval") +
        ylab ("Mean Steps")

```


